---
title: "Explore acoustic telemetry data"
author: "Damiano Oldoni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Explore acoustic telemetry data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

In this vignette we show a typical workflow using the etn package to retrieve **acoustic telemetry data** from ETN. We will also use the packages [dplyr](https://dplyr.tidyverse.org/) for data exploration, [lubridate](https://lubridate.tidyverse.org/) to handle datetime data and [leaflet](https://rstudio.github.io/leaflet/) for interactive visualizations.

To start, load the etn package:

```{r load_pkg}
library(etn)
```

And the other packages (with installation if necessary):

```{r load_other_pkgs, message=FALSE}
other_pkgs <- c("dplyr", "lubridate", "leaflet")

# install missing packages
pkgs_to_install <- other_pkgs[!other_pkgs %in% rownames(installed.packages())]
install.packages(pkgs_to_install)

# load packages
library(dplyr)
library(lubridate)
library(leaflet)
```

Connect with your user credentials (as received by VLIZ) to the database. To not expose such confidential information, you could opt to [use environment variables](https://db.rstudio.com/best-practices/managing-credentials/#use-environment-variables):

```{r connect_to_db}
con <- connect_to_etn(Sys.getenv("userid"), Sys.getenv("pwd"))
# This is the default, so you could also use connect_to_etn()
```

Using `con` as variable to store the collection is not mandatory, but it makes your life much easier as `con` is the default value of the argument `connection`, present in every other function of this package.

## Select project(s) of interest

A researcher typically works in the context of one or more _animal projects_. As the project codes are not always easy to remember, let's start by getting an overview of all projects:

```{r get_animal_projects}
all_projects <- get_animal_projects()
```

Show a preview:

```{r preview_first_10_projects}
all_projects %>% head(10)
```

If you know the `project code` already and you are just interested to get more information about it, you can specify it in the `get_animal_projects()` function directly:

```{r get_demer_project_directly}
projects_code <- c("2014_demer", "2015_dijle")
projects_study <- get_animal_projects(animal_project_code = projects_code)
projects_study
```

This is exactly the same as retrieving all projects first and filtering them afterwards based on column `project_code`:

```{r retrieve_all_first_filter_after}
all_projects %>%
  filter(project_code %in% c(projects_code))
```

To list all available animal project codes as a vector, you can use `list_animal_project_codes()`, one of the etn functions in the [`list_*` family](https://inbo.github.io/etn/reference/index.html#section-list-parameter-options):

```{r list_animal_project_codes}
list_animal_project_codes() %>% head(10)
```

## Animals

By using `get_animals()` you can retrieve information about each animal (`animal_id`), such as scientific name, length, capture/release date and location, and the attached tag(s) (`tag_serial_number`):

```{r get_animals}
animals <- get_animals(animal_project_code = projects_code)
animals %>% head(10)
```

What species and how many individuals are tracked for the projects `2014_demer` and `2015_dijle`?

```{r get_species_and_number_individuals}
animals %>%
  count(scientific_name)
```

## Detections

Let's say we are interested in the tracking data of Wels catfish (_Silurus glanis_) in 2014. You can retrieve the detection history using `get_acoustic_detections()`:

```{r get_detections_silurus}
detections_silurus <- get_acoustic_detections(
  animal_project_code = projects_code,
  start_date = "2014-01-01",
  end_date = "2015-01-01", # The end date is exclusive
  scientific_name = "Silurus glanis"
)
```

Preview:

```{r preview_detections_silurus}
detections_silurus %>% head(10)
```

Which individuals have been detected (`animal_id`) and in which period?

```{r detections_silurus_period}
detections_silurus_period <-
  detections_silurus %>%
  mutate(date = date(date_time)) %>%
  group_by(animal_id) %>%
  summarize(
    start = min(date),
    end = max(date)
  )
detections_silurus_period
```

Notice we group by `animal_id`, the unique identifier of the fish. However, if the fish has only been tagged  once (as typically occurs), we could use `acoustic_tag_id` as well, i.e. the identifier picked up by acoustic receivers:

```{r group_by_acoustic_tag_id}
detections_silurus %>%
  mutate(date = date(date_time)) %>%
  group_by(acoustic_tag_id) %>%
  summarize(
    start = min(date),
    end = max(date)
  )
```

We can also get the tracking duration of each fish:

```{r duration_tracking}
detections_silurus_duration <-
  detections_silurus %>%
  group_by(animal_id) %>%
  summarize(duration = max(date_time) - min(date_time))
detections_silurus_duration
```

How many times has an individual has been detected?

```{r n_detections}
detections_silurus %>%
  group_by(animal_id) %>%
  count()
```

## Stations

At how many detection stations have the individuals been detected?

```{r n_stations}
detections_silurus %>%
  group_by(animal_id) %>%
  distinct(station_name) %>%
  count()
```

Which stations have been involved? You can retrieve them using `list_values` function applied to column `station_name`:

```{r stations_silurus}
stations_silurus <-
  detections_silurus %>%
  list_values(station_name)
stations_silurus
```

Notice how a detection station can be linked to multiple deployments:

```{r link_stations_deployments}
detections_silurus %>%
  distinct(station_name, deployment_id) %>%
  group_by(station_name) %>%
  add_tally() %>%
  arrange(desc(n))
```

Sometimes it's interesting to know the number of detections per station:

```{r n_detections_per_deployment}
n_detect_station <-
  detections_silurus %>%
  group_by(station_name) %>%
  count()
n_detect_station
```

It's also interesting to know the number of unique individuals per station:

```{r n_silurus_station}
n_silurus_station <-
  detections_silurus %>%
  distinct(station_name, animal_id) %>%
  group_by(station_name) %>%
  count()
n_silurus_station
```

## Acoustic tags 

To get more information about the tags involved in `detections_silurus`, you can use the function `get_tags`, which returns tag related information such as serial number, manufacturer, model, and frequency:

```{r acoustic_tags_info}
tags_id <- list_values(detections_silurus, acoustic_tag_id)
tags_silurus <- get_tags(acoustic_tag_id = tags_id)
tags_silurus
```

You can also retrieve such information by `tag_serial_number`:

```{r acoustic_tags_info_via serial_number}
tags_serial <- unique(detections_silurus$tag_serial_number)
tags_silurus <- get_tags(tag_serial_number = tags_serial)
tags_silurus
```

However, keep in mind that there is a fundamental difference between the arguments `acoustic_tag_id` and `tag_serial_number`: the `tag_serial_number` identifies the _device_, which could contain multiple tags or sensors and thus multiple `acoustic_tag_id`.

All possible `acoustic_tag_id` can be retrieved with the correspondent `list_*` function:

```{r list_acoustic_tag_ids}
list_acoustic_tag_ids() %>% head(10)
```

## Acoustic network projects

The detection of Wels catfishes has been possible thanks to one or more acoustic network projects, mentioned in field `acoustic_project_code`. You can retrieve them via the _list_ function `list_values()`:

```{r acoustic_project_codes_involved}
acoustic_project_codes <- detections_silurus %>%
  list_values(acoustic_project_code)
acoustic_project_codes
```

To get more information about these acoustic networks, you can use function `get_acoustic_projects()`

```{r get_acoustic_projects}
acoustic_projects_silurus <- get_acoustic_projects(
  acoustic_project_code = acoustic_project_codes
)
acoustic_projects_silurus
```

You can retrieve the full list of acoustic network projects with the correspondent `list_*` function:

```{r list_acoustic_project_codes}
list_acoustic_project_codes() %>% head(10)
```

## Acoustic deployments

You can retrieve deployment information related to the acoustic networks in `acoustic_project_codes` by using `get_acoustic_deployments()` function:

```{r get_acoustic_deployments_info_networks}
deployments <- get_acoustic_deployments(
  acoustic_project_code = acoustic_project_codes
)
deployments
```

These are the deployments of the acoustic receivers involved in `detections_silurus`:

```{r deploys_silurus}
deploys_silurus <-
  detections_silurus %>%
  list_values(deployment_id)
deploys_silurus
```

More information about them can be retrieved via `get_acoustic_deployments()` function with argument `deployment_id`:

```{r get_acoustic_deployments_info}
deployments_silurus <- get_acoustic_deployments(
  deployment_id = deploys_silurus
)
deployments_silurus
```

Deployment duration:

```{r deployment_duration}
deployments_silurus_duration <-
  deployments_silurus %>%
  mutate(duration = as.duration(recover_date_time - deploy_date_time)) %>%
  select(deployment_id, station_name, duration) %>%
  arrange(deployment_id)
deployments_silurus_duration
```

Number of days a deployment detected the passage of one or more individuals:

```{r n_active_days_deployments_silurus}
n_active_days_deployments_silurus <-
  detections_silurus %>%
  mutate(date = date(date_time)) %>%
  distinct(deployment_id, station_name, date) %>%
  group_by(deployment_id, station_name) %>%
  summarize(n_days = n(), .groups = "drop") %>%
  ungroup()
n_active_days_deployments_silurus
```

Relative detection duration, i.e. number of days with at least one detection divided by deployment duration:

```{r relative_detection_duration}
rel_det_duration_silurus <-
  n_active_days_deployments_silurus %>%
  left_join(
    deployments_silurus_duration,
    by = c("deployment_id", "station_name")
  ) %>%
  mutate(
    relative_detection_duration = n_days * (24 * 60 * 60) / as.numeric(duration)
  ) %>%
  select(deployment_id, station_name, relative_detection_duration)
rel_det_duration_silurus
```

## Data visualization

Aside standard graphs, the geographical component of telemetry data makes interactive maps particularly useful. The package `leaflet` is quite popular to create such kind of visualizations. 

We can for example create a map of the involved stations showing the station name and the acoustic project code it belongs to as pop-ups. First, we retrieve the coordinates of the stations:

```{r geo_info_stations}
geo_info_stations <-
  detections_silurus %>%
  distinct(
    station_name,
    deploy_latitude,
    deploy_longitude,
    acoustic_project_code
  ) %>%
  arrange(station_name)
geo_info_stations
```

To be able to produce the desired map:

```{r map_stations}
leaflet(geo_info_stations) %>%
  addTiles() %>%
  addMarkers(
    lng = ~deploy_longitude,
    lat = ~deploy_latitude,
    popup = ~paste0("Station: ", station_name, " (", acoustic_project_code, ")")
  )
```

We can visualize the number of detections per station, `n_detect_station`, by joining it with `geo_info_stations`:

```{r leaflet_n_detections_Wels_catfish}
# Create a continuous colour palette function
pal <- colorNumeric(
  palette = "viridis",
  domain = n_detect_station$n
)

n_detect_station %>%
  left_join(
    geo_info_stations,
    by = "station_name"
  ) %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~deploy_longitude,
    lat = ~deploy_latitude,
    radius = ~log(n),
    color = ~pal(n),
    fillOpacity = 0.8,
    stroke = FALSE,
    popup = ~paste(
      sep = "<br/>",
      paste0("Station: ", station_name, " (", acoustic_project_code, ")"),
      paste0("# detections: ", n)
    )
  ) %>%
  addLegend(
    title = "Detections",
    pal = pal,
    values = ~n
  )
```

In a similar way we can visualize the number of detected individuals per station, `n_silurus_station`:

```{r map_n_individuals_per_station}
# Create a continuous colour palette function
pal <- colorNumeric(
  palette = "viridis",
  domain = n_silurus_station$n
)

n_silurus_station %>%
  left_join(
    geo_info_stations,
    by = "station_name"
  ) %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~deploy_longitude,
    lat = ~deploy_latitude,
    radius = ~n,
    color = ~pal(n),
    fillOpacity = 0.8,
    stroke = FALSE,
    popup = ~paste(
      sep = "<br/>",
      paste0("Station: ", station_name, " (", acoustic_project_code, ")"),
      paste0("# detected individuals: ", n)
    )
  ) %>%
  addLegend(
    title = "Detected individuals",
    pal = pal,
    values = ~n
  )
```

We can also make a map of the relative detection duration of the deployments. First, we have to retrieve the deployment geographical coordinates:

```{r geo_info_deployments}
geo_info_deploys <-
  detections_silurus %>%
  distinct(
    deployment_id,
    deploy_latitude,
    deploy_longitude,
    station_name,
    acoustic_project_code
  ) %>%
  arrange(station_name)
geo_info_deploys
```

We are ready to create the desired map, where we show the deployment ID, the station name and the network project as popups:

```{r map_rel_detection_duration_deploys}
# Create a continuous colour palette function
pal <- colorNumeric(
  palette = "viridis",
  domain = rel_det_duration_silurus$relative_detection_duration
)

rel_det_duration_silurus %>%
  left_join(
    geo_info_deploys,
    by = c("deployment_id", "station_name")
  ) %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(
    lng = ~deploy_longitude,
    lat = ~deploy_latitude,
    radius = ~100 * relative_detection_duration,
    color = ~pal(relative_detection_duration),
    fillOpacity = 0.8,
    stroke = FALSE,
    clusterOptions = markerClusterOptions(),
    popup = ~paste(
      sep = "<br/>",
      paste0("DeploymentID: ", deployment_id),
      paste0("Station: ", station_name, " (", acoustic_project_code, ")"),
      paste0("# relative detection duration: ", round(relative_detection_duration, 2))
    )
  ) %>%
  addLegend(
    title = "Relative detection duration",
    pal = pal,
    values = ~relative_detection_duration
  )
```

Do you want to add the temporal component to visualize the fish movement? Take a look at the [moveVis](https://movevis.org/index.html) package.
