---
title: "Configure credentials"
output: rmarkdown::html_vignette
author: Pieter Huybrechts
date: "2025-07-22"
vignette: >
  %\VignetteIndexEntry{Configure credentials}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

All etn functions require credentials (username and password) to connect to the ETN database.

## Register for an ETN account

Don't have an ETN user account? [Register here](https://rshiny.lifewatch.be/account?p=register).

## Storing credentials

By default, all functions will ask you for your credentials. Very tedious! Avoid this by storing your credentials (once) in a local `.Renviron` file and etn will use those rather than asking every time.

1. Open your `.Renviron` file with:

    ```{r, eval=FALSE}
    # install.packages(usethis)
    usethis::edit_r_environ()
    ```

2. Add the following lines to the file and save:

    ```bash
    ETN_USER = "your username"
    ETN_PWD = "your password"
    ```

3. Restart R.
4. Try:

   ```{r, eval=FALSE}
   library(etn)
   get_animal_projects() # This should return a data frame
   ```

## Forgot your password?

Use the [reset password form](https://rshiny.lifewatch.be/account?p=lostpass).
