---
title: "Explore ETN acoustic telemetry data"
author: "Damiano Oldoni"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Explore ETN data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

In this vignette we show a typical workflow using `etn` package to retrieve acoustic telemetry data from the postgreSQL ETN database and the [`dplyr`](https://dplyr.tidyverse.org/) package (part of the [tidyverse](https://www.tidyverse.org/) R pacakges collection)  to perform basic data exploration. We will also use ome other packages: [`lubridate`](https://lubridate.tidyverse.org/) to handle date-time data and [`leaflet`](https://rstudio.github.io/leaflet/) and [`htmltools`](https://rstudio.github.io/htmltools/index.html) for interactive visualizations.

To start, load the `etn`package:

```{r load_pkg}
library(etn)
```

and the other packages (install first the ones missing):

```{r load_other_pkgs, message=FALSE}
other_pkgs <- c("dplyr", "lubridate", "leaflet", "htmltools")

# install missing packages
pkgs_to_install <- other_pkgs[!other_pkgs %in% rownames(installed.packages())]
install.packages(pkgs_to_install)

# load packages
library(dplyr)  #to do datascience
library(lubridate)  #to handle date-time data
library(leaflet)  # to create interactive maps
library(htmltools)  # to write HTML in R
```

Connect with your user credentials (as received by VLIZ) to the database. To not expose such confidential information, you could opt to [use environment variables](https://db.rstudio.com/best-practices/managing-credentials/#use-environment-variables):

```{r connect_to_db}
con <- connect_to_etn(Sys.getenv("userid"), Sys.getenv("pwd"))
```

Using `con` as variable to store the collection is not mandatory, but it makes your life much easier as `con` is the default value of the argument `connection`, present in every other function of this package.

## Select project(s) of interest

A researcher typically works in the context of one or more _animal projects_. As the project names are not easy to remember, it is always good to start by an overview of all projects we have access to:

```{r get_my_animal_projects}
all_my_projects <- get_animal_projects()
```

Show a preview:

```{r preview_first_10_projects}
# show the first 10 projects
all_my_projects %>% head(10)
```

If you know the `project code` already and you are just interested to get more information about it, you can specify it in the `get_animal_projects()` function directly:

```{r get_demer_project_directly}
projects_code <- c("2014_demer", "2015_dijle")
projects_study <- get_animal_projects(animal_project_code = projects_code)
projects_study
```

This is exactly the same as retrieving all projects first and filtering them afterwards based on column `project_code`:

```{r retrieve_all_first_filter_after}
all_my_projects %>%
  filter(project_code %in% c(projects_code))
```

If you are just interested in the project names you have right to access, you can easily retrieve them as a vector via `list_animal_project_codes`, one of the functions of the `list_*` family:

```{r list_animal_project_codes}
list_animal_project_codes()
```

## Data exploration

### Animals under study

By using `get_animals()` you can retrieve information about each fish (`animal_id`), such as scientific name and length, capture/release date and location and the atttached acoustic transmitter (_tag_):

```{r get_animals}
animals <- get_animals(animal_project_code = projects_code)
# preview
animals %>% head(10)
```

Which species and how many individuals have been tracked for these projects?

```{r get_species_and_number_individuals}
animals %>%
  count(scientific_name)
```

### Detection history

Let's say we are interested in the tracking data of Wels catfish (_Silurus glanis_) between begin 2014 and end 2015:

```{r get_detections_silurus}
detections_silurus <- get_acoustic_detections(
  animal_project_code = projects_code,
  start_date = "2014-01-01",
  end_date = "2015-12-31",
  scientific_name = "Silurus glanis")
```

Preview:

```{r preview_detections_silurus}
detections_silurus %>% head(10)
```

Which fishes have been detected (`animal_id`) and in which period?

```{r detections_silurus_period}
detections_silurus_period <-
  detections_silurus %>%
  mutate(date = date(date_time)) %>%
  group_by(animal_id) %>%
  summarise(start = min(date),
            end = max(date))
detections_silurus_period
```

Notice we group by `animal_id`, the unique identifier of the fish. However, if the fish has been tagged only once (as typically occurs), we could use the `acoustic_tag_id`, i.e. the ID of the transmitter, an idea the researcher has more feeling with:

```{r group_by_acoustic_tag_id}
detections_silurus %>%
  mutate(date = date(date_time)) %>%
  group_by(acoustic_tag_id) %>%
  summarise(start = min(date),
            end = max(date))
```

We can also get the tracking duration of each fish:

```{r duration_tracking}
detections_silurus_duration <-
  detections_silurus %>%
  group_by(animal_id) %>%
  summarise(duration = max(date_time) - min(date_time))
detections_silurus_duration
```

How many detections per fish?

```{r n_detections}
detections_silurus %>%
  group_by(animal_id) %>%
  count()
```

### Detection stations

At how many detection stations have the fishes been detected?

```{r n_stations}
detections_silurus %>%
  group_by(animal_id) %>%
  distinct(station_name) %>%
  count()
```

Which stations have been involved?

```{r deployments_silurus}
stations_silurus <-
  detections_silurus %>%
  select(station_name) %>%
  distinct() %>%
  arrange()
stations_silurus
```

Notice how a detection station can be linked to multiple deployments:

```{r link_stations_deployments}
detections_silurus %>%
  distinct(station_name, deployment_id) %>%
  group_by(station_name) %>%
  add_tally() %>%
  arrange(desc(n))
```

Sometimes it's interesting to know the number of detections per station:

```{r n_detections_per_deployment}
n_detect_station <- 
  detections_silurus %>%
  group_by(station_name) %>%
  count()
n_detect_station
```

It's also interesting to know the number of unique detected fishes per station:

```{r n_silurus_station}
n_silurus_station <- 
  detections_silurus %>%
  distinct(station_name, animal_id) %>%
  group_by(station_name) %>%
  count()
n_silurus_station
```


### Acoustic tags 

To get more information about the tags involved in `detections_silurus`, you can use the function `get_tags`, which returns information such as manufacturer, model and frequency:

```{r acoustic_tags_info}
tags_id <- unique(detections_silurus$acoustic_tag_id)
tags_silurus <- get_tags(acoustic_tag_id = tags_id)
tags_silurus
```

You can also retrieve such information by `tag_serial_number`

```{r acoustic_tags_info}
tags_serial <- unique(detections_silurus$tag_serial_number)
tags_silurus <- get_tags(tag_serial_number = tags_serial)
tags_silurus
```

However, keep in mind that there is a fundamental differnce between these two arguments: the `tag_serial_number` identifies the instrument, which could contain multiple tags. Not only, the argument `acoustic_tag_id` is only valid when you are dealing with acoustic telemetry data (`get_acoustic_*()` functions).

### Acoustic deployments

These are the deployments of the acoustic receivers involved in `detections_silurus`:

```{r deploys_silurus}
deploys_silurus <- 
  detections_silurus %>%
  distinct(receiver_id, deployment_id)
deploys_silurus
```

More information about them can be retrieved by `get_acoustic_deployments()` function:

```{r get_acoustic_deployments_info}
deployments_silurus <-get_acoustic_deployments(
  receiver_id = deploys_silurus$receiver_id
)
```

Deployment duration:

```{r deployment_duration}
deployments_silurus_duration <- 
  deployments_silurus %>%
  mutate(duration = as.duration(recover_date_time - deploy_date_time)) %>%
  select(deployment_id, station_name, duration)
deployments_silurus_duration
```

Number of days a deployment detected the passage of one or more fishes:

```{r n_active_days_deployments_silurus}
n_active_days_deployments_silurus <- 
  detections_silurus %>%
  mutate(date = date(date_time)) %>%
  distinct(deployment_id, station_name, date) %>%
  group_by(deployment_id, station_name) %>%
  summarise(n_days = n()) %>%
  ungroup()
n_active_days_deployments_silurus
```

Relative detection duration, i.e. number of days with at least one detection divided by deployment duration:

```{r relative_detection_duration}
rel_det_duration_silurus <- 
  n_active_days_deployments_silurus %>%
  left_join(deployments_silurus_duration,
            by = c("deployment_id", "station_name")) %>%
  mutate(relative_detection_duration = n_days*(24*60*60) / as.numeric(duration)) %>%
  select(deployment_id, station_name, relative_detection_duration)
rel_det_duration_silurus
```

### Data visualization

The package `leaflet` is very useful to create interactive maps. We start creating a map of the involved sattions showing the station name and the acoustic project code it belongs to as pop-up. First, we have to retrieve the station geographical coordinates:

```{r geo_info_stations}
geo_info_stations <- 
  deployments_silurus %>%
  distinct(station_name, 
          deploy_latitude, 
          deploy_longitude, 
          acoustic_project_code)
geo_info_stations
```

And then we add them on a map as markers: 

```{r map_stations}
leaflet(geo_info_stations) %>%
  addTiles() %>%
  addMarkers(lng = ~deploy_longitude,
             lat = ~deploy_latitude,
             popup = ~htmlEscape(
               paste0("Station: ", station_name, "\n",
                      "(", acoustic_project_code, ")")))
```

In the previous section we calculated the number of detections per station, `n_detect_station`. We can visualize such information on a map by joining it with `geo_info_stations`:

```{r leaflet_n_detections_Wels_catfish}
# Create a continuous palette function
pal <- colorNumeric(
  palette = "viridis",
  domain = n_detect_station$n)

n_detect_station %>%
  left_join(geo_info_stations,
              by = "station_name") %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lng = ~deploy_longitude,
             lat = ~deploy_latitude,
             radius = ~log(n),
             color = ~pal(n),
             fillOpacity = 0.8,
             stroke = FALSE,
             popup = ~paste(sep = "<br/>",
                            paste0("Station: ", 
                                   station_name, 
                                   "(", acoustic_project_code, ")"),
                            paste0("# detections: ", n))) %>% 
  addLegend(title = "Number of detections",
            pal = pal,
            values = ~n)
```

On a similar way we can visualize the number of detected fishes per station, `n_silurus_station`:

```{r map_n_fishes_per_station}
# Create a continuous palette function
pal <- colorNumeric(
  palette = "viridis",
  domain = n_silurus_station$n)

n_silurus_station %>%
  left_join(geo_info_stations,
              by = "station_name") %>%
  leaflet() %>%
  addTiles() %>%
  addCircleMarkers(lng = ~deploy_longitude,
             lat = ~deploy_latitude,
             radius = ~n,
             color = ~pal(n),
             fillOpacity = 0.8,
             stroke = FALSE,
             popup = ~paste(sep = "<br/>",
                            paste0("Station: ", 
                                   station_name, 
                                   "(", acoustic_project_code, ")"),
                            paste0("# detected fishes: ", n))) %>% 
  addLegend(title = "Number of detections",
            pal = pal,
            values = ~n)
```

