---
title: "Introduction to main functions of the ETN package"
author: "Stijn Van Hoey"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    df_print: paged
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Introduction

The `etn` package provides a set of `get_*`-based functions to access the ETN data and metadata. In this vignette a quick overview of these main functions is given. As the access to data is linked to specific user profiles, keep in mind that the results can be different for other users.

To start, load the package:

```{r}
library(etn)
```

When not installed, do not forget to install the package using the `devtools` package:

```{r eval=FALSE, echo=TRUE}
# install.packages("devtools")
devtools::install_github("inbo/etn")
```

## The database connection

In order to access the data, the first step is always to connect with your user credentials (as received by VLIZ) to the database. 

```{r}
my_con <- connect_to_etn(Sys.getenv("userid"),
                         Sys.getenv("pwd"))
my_con
```

Remark that for this example, we [load the user credentials as environmental variables](http://db.rstudio.com/best-practices/managing-credentials/#use-environment-variables) by saving an `.Renviron` file in the home directory with the username and password stored. Other options are available as well, as explained [here](http://db.rstudio.com/best-practices/managing-credentials/).

Once the connection established, the `get_*` functions can be used to effectively access the data. Each of these functions have a (valid database) connection as first input.

Notice, that you can find the `get_*`-functions inside Rstudio by typing: `etn::get_*` + TAB button.

## Metadata

### Project overview

To get a full overview of the projects you have access to, use the function `get_projects()`:

```{r projects}
my_projects <- get_projects(my_con)
head(my_projects, 10) # print 10 first
```

As projects are defined both on the level of the `animal` as well as the `network`, a column `type` is available that defines the type of project (`animal` or `network`). Still, one can also directly subset by adding the type as input argument, e.g. for network:

```{r my_animal_projects}
my_network_projects <- get_projects(my_con, project_type = "network")
head(my_network_projects, 10) # print 10 first
```

### Receiver overview

A network project typically considers multiple receivers that are placed on specific locations to lay out the network. To get an overview of the receivers or those linked to specific projects, use the `get_receivers` function:

```{r}
ws3_receivers <- get_receivers(my_con, network_project = "ws3")
head(ws3_receivers, 10) # print 10 first
```

You can also define multiple (network) projects at the same time, by combining them in a vector:

```{r}
ws_receivers <- get_receivers(my_con, network_project = c("ws1", "ws3"))
head(ws_receivers, 10) # print 10 first
```

When a number of columns is not useful to your specific purpose, remember that the output is a `data.frame`, which can be filtered and subsetted conveniently with for example the [dplyr](http://dplyr.tidyverse.org/) functions. 

```{r}
library(dplyr)
```

For example, selecting only the columns `serial_number`, `receiver` and `status`:

```{r}
ws_receivers %>% 
  select(serial_number, receiver, status) %>%
  head(10) # print 10 first
```

### Deployments overview

Receivers are deployed for a specific period and this deployment information can be accessed as well with the command `get_deployments`. Filtering on the network project is similar to `get_receivers` and as the information about the receiver `status` is crucial, the function provides an additional argument to query based on the receiver status. Getting an overview of the lost and broken receivers of a specific project:

```{r}
my_deployments <- get_deployments(my_con, network_project = "ws1", 
                                  receiver_status = c("Broken", "Lost"))
head(my_deployments, 10) # print 10 first 
```

### Animals overview

Within an animal project, multiple animals are provided with a transmitter. The metadata of these animals is requested with the function `get_animals`. As animals can be part of both network and animal projects, both can be used to filter the results. 

As such, selecting the animals metadata for the `2012_leopoldkanaal` project:

```{r}
my_animals <- get_animals(my_con, animal_project = "2012_leopoldkanaal")
head(my_animals, 10) # print 10 first
```

When interested in specific species, add the scientific name as argument.

```{r}
my_animals <- get_animals(my_con, animal_project = "phd_reubens",
                       scientific_name = "Sentinel")
head(my_animals, 10) # print 10 first
```

Remark: currently these names are NOT yet verified with a taxonomic backbone as this is not yet done for the names of the database managed by VLIZ. Support will be added when the database names are as well.

### Transmitter overview

The transmitters/tags metadata can be accessed by the `get_transmitters` function, with the option to subset on the animal project names:

```{r}
my_tags <- get_transmitters(my_con, animal_project = "phd_reubens")
head(my_tags, 10) # print 10 first
```

## Detections data

The raw detections are the core for any type of analysis and access is provided by the `get_detections` function. Different filters are available which you can explore by typing `?get_detections` in the console. 

To narrow down the data request, filters are provided on animal/network project name, the deployment station name and the transmitter identifier to narrow. Furthermore, as these data sets can be large, some additional time-based filters are provided with a `start_date` and an `end_date`: 

```{r}
my_detections <- get_detections(my_con, animal_project = "phd_reubens", 
                                start_date = "2011-01-28", end_date = "2011-02-01")
head(my_detections, 10) # print 10 first
```

The date level can be defined on month (e.g. `"2011-01"`) or year (e.g. `"2011"`) level as well. Finally, for explorative reasons, a hard maximum limit of records to request can help to have a quick glance on the expected data.

See more exampled in the function help (remember the `?get_detections`).
