---
title: "ETN field definitions"
author: "Peter Desmet & Filip Waumans"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ETN field definitions}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<p class="alert alert-warning">
The information below is outdated. We are working on [an update](https://github.com/inbo/etn/issues/226).
</p>

```{r include=FALSE, eval=FALSE}
library(etn)
library(readr)
library(dplyr)
library(tidyr)
library(stringr)
library(glue)
library(DBI)
library(kableExtra)
```

```{r include=FALSE, eval=FALSE}
con <- connect_to_etn()
query <- glue_sql("
  SELECT
    data_format_field_views_definition AS view,
    lower(field_name) AS field,
    trim(regexp_replace(field_definition, '[\\s+]+', ' ', 'g')) AS definition,
    trim(regexp_replace(field_units_format_description, '[\\s+]+', ' ', 'g')) AS example,
    view_order AS \"order\"
  FROM
    app.field_metadata
  WHERE
    data_format_field_views_definition IS NOT NULL
  ORDER BY
    view,
    \"order\"
  ", .con = connection
)
etn_fields <- dbGetQuery(con, query)
```

```{r include=FALSE, eval=FALSE}
write_csv(etn_fields, "../inst/assets/etn_fields.csv", na = "")
```

Field definitions are managed in the ETN database, not in this package, but suggestions/corrections can be reported as an issue (see the [contributing guide](../.github/CONTRIBUTING.md)). The definitions are also available as a [csv file](https://github.com/inbo/etn/blob/master/inst/assets/etn_fields.csv).

```{r include=FALSE, eval=FALSE}
# Create helper function to display tables for each view
view_table <- function(view_name) {
  view_data <-
    etn_fields %>%
    # Filter on view
    filter(view == view_name) %>%
    # Order by order (should already be done by query)
    arrange(.data$order) %>%
    # Select columns
    select(field, definition, example) %>%
    # Convert quotes to backticks, so examples are shown as markdown monospaced code
    mutate(across(everything(), ~ str_replace_all(., "\"", "`"))) %>%
    # Wrap "field" in backticks
    mutate(field = if_else(!is.na(field), paste0("`", field, "`"), field)) %>%
    # Convert "NA" to empty strings
    mutate(across(everything(), ~ replace_na(., "")))
  
  view_kable <-
    kable(view_data) %>%
    # field: some have long names, break-word will force them to wrap
    column_spec(1, width = "25%", extra_css = "word-break: break-word") %>%
    # definition
    column_spec(2, width = "40%") %>%
    # example
    column_spec(3, width = "35%")
  
  return(view_kable)
}
```

## Animals

Related function: `get_animals()`.

```{r echo=FALSE, eval=FALSE}
view_table("animals_view2")
```

## Tags

Related function: `get_tags()`.

```{r echo=FALSE, eval=FALSE}
view_table("tags_view2")
```

## Detections

Related function: `get_detections()`.

```{r echo=FALSE, eval=FALSE}
view_table("detections_view2")
```

## Deployments

Related function: `get_deployments()`.

```{r echo=FALSE, eval=FALSE}
view_table("deployments_view2")
```

## Receivers

Related function: `get_receivers()`.

```{r echo=FALSE, eval=FALSE}
view_table("receivers_view2")
```
