---
title: "Package options"
output: rmarkdown::html_vignette
author: Pieter Huybrechts
date: "2025-09-26"
vignette: >
  %\VignetteIndexEntry{Package options}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A number of etn options can be configured as [Environmental Variables](https://stat.ethz.ch/R-manual/R-devel/library/base/html/EnvVar.html).

## Setting options

To temporarily set an option for the duration of an R session, use:

```{r, eval=FALSE}
Sys.setenv(ETN_PROTOCOL = "opencpu")
# Retrieve with Sys.getenv("ETN_PROTOCOL")
```

To store an option permanently, open the `.Renviron` file with `usethis::edit_r_environ()`, edit, and save:

```bash
ETN_PROTOCOL = "opencpu"
```

## Available options

- `ETN_USER`: Your username to the ETN database. See [authentication](etn_credentials.html).

- `ETN_PWD`: Your password to the ETN database. See [authentication](etn_credentials.html).

- `ETN_PROTOCOL` (value: `"opencpu"` or `"localdb"`): etn will automatically detect how to best connect to the ETN database, i.e. it will use a local database connection when available. This option allows you to override this behaviour by hardcoding the protocol to use.

- `ETN_TEST_API` (value: a url) **developers only**: etn uses the production version of the OpenCPU API by default. This option allows you to use a test version of the OpenCPU API (available upon request). Any other value will break the package.
