---
title: "Darwin Core mapping for occurrence dataset"
author: Peter Desmet
date: "`r Sys.Date()`"
output:
  md_document:
    variant: markdown_github
    pandoc_args: ["--atx-headers"]
---

# `r rmarkdown::metadata$title`

By: `r rmarkdown::metadata$author`

Date: `r rmarkdown::metadata$date`

```{r configure_knitr, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

## Setup

Set locale (so we use UTF-8 character encoding):

```{r}
# This works on Mac OS X, might not work on other OS
Sys.setlocale("LC_CTYPE", "en_US.UTF-8")
```

Load libraries:

```{r}
library(tidyverse) # For data transformations

# None core tidyverse packages:
library(magrittr)  # For %<>% pipes

# Other packages
library(janitor)   # For cleaning input data
library(knitr)     # For nicer (kable) tables
```

Set file paths (all paths should be relative to this script):

```{r}
raw_data_file = "../data/raw/denormalized_observations_50000.csv"
processed_dwc_occurrence_file = "..data/processed/dwc_occurrence/occurrence.csv"
```

## Read data

Read the source data:

```{r}
raw_data <- read.csv(raw_data_file, fileEncoding = "UTF-8")
```

Clean data somewhat:

```{r}
raw_data %>%
  # Remove empty rows
  remove_empty_rows() %>%
  # Have sensible (lowercase) column names
  clean_names() -> raw_data
```

Add prefix `raw_` to all column names to avoid name clashes with Darwin Core terms:

```{r}
colnames(raw_data) <- paste0("raw_", colnames(raw_data))
```

Save those column names as a list (makes it easier to remove them all later):

```{r}
raw_colnames <- colnames(raw_data)
```

Preview data:

```{r}
kable(head(raw_data))
```

## Create occurrence core

### Pre-processing

```{r}
occurrence <- raw_data
```

Sort by transmitter and date:

```{r}
occurrence %<>% arrange(raw_transmitter, raw_datetime)
```

### Term mapping

Map the source data to [Darwin Core Occurrence](http://rs.gbif.org/core/dwc_occurrence_2015-07-02.xml) (but in the classic Darwin Core order):

#### type

```{r}
occurrence %<>% mutate(type = "Event")
```

#### modified
#### language

```{r}
occurrence %<>% mutate(language = "en")
```

#### license

```{r}
occurrence %<>% mutate(license = "http://creativecommons.org/publicdomain/zero/1.0/")
```

#### rightsHolder
#### accessRights

```{r}
occurrence %<>% mutate(accessRights = "http://www.inbo.be/en/norms-for-data-use")
```

#### bibliographicCitation
#### references
#### institutionID
#### collectionID
#### datasetID
#### institutionCode
#### collectionCode
#### datasetName
#### ownerInstitutionCode
#### basisOfRecord
#### informationWithheld
#### dataGeneralizations
#### dynamicProperties
#### occurrenceID
#### catalogNumber
#### recordNumber
#### recordedBy
#### individualCount
#### organismQuantity
#### organismQuantityType
#### sex
#### lifeStage
#### reproductiveCondition
#### behavior
#### establishmentMeans
#### occurrenceStatus
#### preparations
#### disposition
#### associatedMedia
#### associatedReferences
#### associatedSequences
#### associatedTaxa
#### otherCatalogNumbers
#### occurrenceRemarks
#### organismID
#### organismName
#### organismScope
#### associatedOccurrences
#### associatedOrganisms
#### previousIdentifications
#### organismRemarks
#### materialSampleID
#### eventID
#### parentEventID
#### fieldNumber
#### eventDate
#### eventTime
#### startDayOfYear
#### endDayOfYear
#### year
#### month
#### day
#### verbatimEventDate
#### habitat
#### samplingProtocol
#### sampleSizeValue
#### sampleSizeUnit
#### samplingEffort
#### fieldNotes
#### eventRemarks
#### locationID
#### higherGeographyID
#### higherGeography
#### continent
#### waterBody
#### islandGroup
#### island
#### country
#### countryCode
#### stateProvince
#### county
#### municipality
#### locality
#### verbatimLocality
#### minimumElevationInMeters
#### maximumElevationInMeters
#### verbatimElevation
#### minimumDepthInMeters
#### maximumDepthInMeters
#### verbatimDepth
#### minimumDistanceAboveSurfaceInMeters
#### maximumDistanceAboveSurfaceInMeters
#### locationAccordingTo
#### locationRemarks
#### decimalLatitude
#### decimalLongitude
#### geodeticDatum
#### coordinateUncertaintyInMeters
#### coordinatePrecision
#### pointRadiusSpatialFit
#### verbatimCoordinates
#### verbatimLatitude
#### verbatimLongitude
#### verbatimCoordinateSystem
#### verbatimSRS
#### footprintWKT
#### footprintSRS
#### footprintSpatialFit
#### georeferencedBy
#### georeferencedDate
#### georeferenceProtocol
#### georeferenceSources
#### georeferenceVerificationStatus
#### georeferenceRemarks
#### geologicalContextID
#### earliestEonOrLowestEonothem
#### latestEonOrHighestEonothem
#### earliestEraOrLowestErathem
#### latestEraOrHighestErathem
#### earliestPeriodOrLowestSystem
#### latestPeriodOrHighestSystem
#### earliestEpochOrLowestSeries
#### latestEpochOrHighestSeries
#### earliestAgeOrLowestStage
#### latestAgeOrHighestStage
#### lowestBiostratigraphicZone
#### highestBiostratigraphicZone
#### lithostratigraphicTerms
#### group
#### formation
#### member
#### bed
#### identificationID
#### identificationQualifier
#### typeStatus
#### identifiedBy
#### dateIdentified
#### identificationReferences
#### identificationVerificationStatus
#### identificationRemarks
#### taxonID
#### scientificNameID
#### acceptedNameUsageID
#### parentNameUsageID
#### originalNameUsageID
#### nameAccordingToID
#### namePublishedInID
#### taxonConceptID
#### scientificName
#### acceptedNameUsage
#### parentNameUsage
#### originalNameUsage
#### nameAccordingTo
#### namePublishedIn
#### namePublishedInYear
#### higherClassification
#### kingdom
#### phylum
#### class
#### order
#### family
#### genus
#### subgenus
#### specificEpithet
#### infraspecificEpithet
#### taxonRank
#### verbatimTaxonRank
#### scientificNameAuthorship
#### vernacularName
#### nomenclaturalCode
#### taxonomicStatus
#### nomenclaturalStatus
#### taxonRemarks

### Post-processing

Remove the original columns:

```{r}
occurrence %<>% select(-one_of(raw_colnames))
```

Preview data:

```{r}
kable(head(occurrence))
```
